Facility and Time Step Discretization Effects in Fuel Cycle Simulation

### 2015-09-14 paper.tex
more introduction wrangling

Traditionally, many fuel cycle simulators use a system dynamics approach
\cite{forrester_industrial_1961} to modeling. These simulators model several
"stocks" that have levels that change over time due to "flows" between them.
The flows are determined by an equation that is a function of the state of the
system (i.e.  the system's stocks and flows).  A modeled system is then solved
in discrete or continuous time depending on the limitations of the software
and constraints imposed by the model.  In system dynamics simulators, the
number of stocks and potential flows between them is generally a static
property of the modeled system, making it more difficult to swap different
reactor models in and out of the same scenario without wiring them into the
simulator manually.

System dynamics based fuel cycle simulators generally use a collection of
stocks to represent the state of groups of like facilities commonly referred
to as "fleets".  As the simulation steps through time, the levels of stocks
(e.g. reactor fresh fuel inventory, repository waste inventories, etc.) are
adjusted according to the calculated flow values.  The VISION simulator
\cite{jacobson_verifiable_2010} among others operates this way.  In
fleet-based modeling:

@@

    \item resource flows are continuous (limited by time step discretization)

@@

Cyclus' agent-based architecture in concert with its dynamic resource exchange
has the ability to model fleet-based facilities.  However, Cyclus was designed
with the intent of modeling facilities individually with discrete flows
enabling the investigation of interesting real-world effects such as
competition, reactor cycle staggering, and individual reactor outages.  The
difference between these two modeling styles in terms of their affects on
output of fuel cycle simulators in general is currently unknown. This analysis
is designed to provide both quantitative and qualitative insight about how
fleet-based continuous flow models differ from individual facility discrete
flow models.

Many fuel cycle simulators step through time in discrete steps.  The duration
of a single time step can have significant impact on the dynamics of a
simulation.  The time step duration can affect the magnitude of individual
material transfers between facilities and correspondingly affects the noise
level in facility material inventories.  It can also act as a minimum bound
for outage durations. The effects of the time step duration will also be
investigated together with fleet vs facility modeling.  Questions of interest
include:

@@

This work investigates in detail differences between fleet vs. individual
reactor modeling together with variations in time step duration.

@@

One important phenomenon related to time step duration is the \emph{inventory
drawdown effect}.  Increasing time step duration reduces the frequency over
which refueling can occur, resulting in larger impulse drawdown on
inventories. This by itself is not problematic because this is balanced by
larger inventory top-up quantities.  Another component of this effect is the
simultaneity of incoming and outgoing material flows on a given time step.
Facilities do not know about incoming material when they make commitments for
outgoing material.  For longer time steps, larger incoming quanta of material
are not available for making offers. In general longer time steps and
individual reactor modeling create a need for larger floating inventory
buffers in order to avoid temporary material shortages.

There are also performance differences between the two modeling paradigms.  A
single 200 year simulation with several hundred fleet-based reactors takes
about 2 seconds to run on a typical desktop machine where the same simulation
using individual reactor facilities takes about 20 seconds.  This order of
magnitude difference is an important tradeoff to keep in mind when doing fuel
cycle analysis in general, particularly when considering optimization or
sensitivity studies.

@@

A fleet reactor facility was created for Cyclus.  It is possible to create
fleet models for Cyclus that live anywhere along the spectrum from individual,
discrete facilities to continuous flow fleet facilities.  The particular model
created here falls more toward the aggregate, continuous flow side of the
spectrum.

>>>>

\subsection{Motivation and Previous Studies}

Due to the diversity of fuel cycle simulators' modeling assumptions direct
comparison and benchmarking can be difficult.  In 2012 the Organisation for
Economic Co-operation and Development (OECD) completed a benchmark study
\cite{oecd2012benchmark} that is perhaps one of the most complete published
comparisons performed.  Despite this, however, various results from the
different simulators were often significantly different due to different
modeling decisions such as reprocessing strategies, refueling behavior,
reactor end-of-life handling, etc.

Traditionally, many fuel cycle simulators have used a system dynamics approach
\cite{forrester_industrial_1961} to modeling. These simulators model several
"stocks" that have levels that change over time due to "flows" between them.
The flows are determined by an equation that is a function of the state of the
system (i.e. the system's stocks).  A constructed system is then solved
in discrete or continuous time depending on the limitations of the software
and constraints imposed by the model.  In system dynamics simulators, the
number of stocks and potential flows between them is generally a static
property of the modeled system, making it more difficult to swap different
reactor models in and out of the same scenario without wiring them into the
simulator manually.

System dynamics based fuel cycle simulators generally use a collection of
stocks to represent the state of groups of like facilities commonly referred
to as "fleets".  As the simulation steps through time, the levels of stocks
(e.g. reactor fresh fuel inventory, repository waste inventories, etc.) are
adjusted according to the calculated flow values.  Simulators that model
facility fleets built on a system dynamics foundation include VISION
\cite{jacobson_verifiable_2010}, DANESS \cite{van_den_durpel_daness_2009},
DYMOND (predecessor to DANESS), and ORION among others.  In fleet-based
modeling:

@@

    \item resource flows are often (approximately) continuous (limited by time
        step discretization)

@@

This work investigates in detail differences between fleet vs. individual
reactor modeling.  The difference between these two modeling styles in terms
of their affects on output of fuel cycle simulators in general is currently
unknown. This analysis is designed to provide both quantitative and
qualitative insight about how fleet-based continuous flow models differ from
individual facility discrete flow models.

Many fuel cycle simulators step through time in discrete steps.  The duration
of a single time step can have significant impact on the dynamics of a
simulation.  The time step duration can affect the magnitude of individual
material transfers between facilities and correspondingly affects the noise
level in facility material inventories.  It can also act as a minimum bound
for outage durations. The effects of the time step duration will also be
investigated together with fleet vs facility modeling.

Questions of interest include:

@@

\subsection{CYCLUS}

Cyclus' agent-based architecture in concert with its dynamic resource exchange
has the ability to model fleet-based facilities.  However, Cyclus was designed
to enable modeling facilities individually with discrete flows enabling the
investigation of interesting real-world effects such as competition, reactor
cycle staggering, and individual reactor outages.

@@

One important phenomenon related to time step duration is the \emph{inventory
drawdown effect}.  Increasing time step duration reduces the frequency over
which refueling can occur, resulting in larger impulse drawdown on
inventories. This by itself is not problematic because this is balanced by
correspondingly larger inventory top-up quantities.  The other component of
this effect is the simultaneity of incoming and outgoing material flows on a
given time step.  Within a particular time step facilities do not know about
incoming material when they make commitments for outgoing material.  For
longer time steps, larger incoming quanta of material are not available for
making offers. In general longer time steps and individual reactor modeling
create a need for larger floating inventory buffers in order to avoid
material shortages.

There are also performance differences between the two modeling paradigms.  A
single 200 year simulation with several hundred fleet-based reactors takes
about 2 seconds to run on a typical desktop machine where the same simulation
using individual reactor facilities takes about 20 seconds.  This order of
magnitude difference is an important tradeoff to keep in mind when doing fuel
cycle analysis in general, particularly when considering optimization or
sensitivity studies where running hundreds of thousands of simulations may be
desirable.

@@

A fleet reactor facility was created for Cyclus.  The flexibility of Cyclus
makes it relatively straight forward to create fleet models for Cyclus that
live anywhere along the spectrum from individual, discrete facilities to
continuous flow fleet facilities.  In order to contrast well with individual,
single-facility granularity modeling, the fleet model created here falls more
toward the aggregate, continuous flow side of the spectrum.
